A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-08-05 docs/introduction_v1.01.tex
Started writing sections on MGE photometry.

\paragraph{Characteristics of J1331}
\begin{itemize}
\item SDSS J1331+3638 (J1331)
\item approximate hubble type Sb
\item first discovered by Sloan digital sky survey (SDSS) [TO DO: REF]
\item at redshift $z_d\simeq0.113$ [TO DO: REF]
\item \citet{SWELLSI} identified it as a strong lens
\item large reddish bulge and bluish spiral arms, see panel (a) and (b) in figure 1
\item superimposed by quadruplet of extended bluish images at a redshift of $z_s\simeq0.254$ [TO DO: REF], see panel (c) in figure 1
\item lensed object might be a star-forming blob of a background galaxy.
\item Lensing properties first analysed by \citet{SWELLSIII}
\item rather edge-on $\longrightarrow$ possible to measure rotation curves. \citet{SWELLSV} measured the gas and stellar rotation curves along the major axis. Fitted galaxy model to gas kinematics at large radii, and lensing result
\item large counter-rotating core, see panel (d) in figure 1
\item possible minor merger in the past
\end{itemize}

@@

\paragraph{Data used}
\begin{itemize}
\item Hubble Space Telescope (HST) imaging by \cite{SWELLSI}, see panels (a) and (b) in Figure 1
\item  \citet{SWELLSV} measured the gas and stellar rotation curves along the major axis. see panels (b) and (d) in Figure 1
\end{itemize}

@@

\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in F450W ("blue")}
  \label{fig:???}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in F814W ("red")}
  \label{fig:???}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{The lensing images}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar Kinematics by \citet{SWELLSV}}
  \label{fig:???}
\end{subfigure}
\caption{Hubble Space telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3638 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The galaxy's coordinates on the sky are right ascension $\alpha$ = 202.91800$^\circ$ and declination $\delta$ = 36.46999$^\circ$ (epoch J2000). Image orientation and scaling are indicated in panel (a); the scaling transformation from arcseconds to the physical size of the galaxy in kpc uses the galaxy's redshift $z_d = 0.113$ \citep{SWELLSIII}. The black solid line in panel (a) shows the orientation of the major-axis. The line has a length of 10 arcsec and indicates the region within which we carry out the Jeans modelling. (NOT ALL THE TIME.???????) \emph{Panel (c):} The central region of J1331 in F450W, surface brightness subtracted. An IRAF ellipse ???? fit to the F450W surface brightness in panel (a) was subtracted from the image. The (smoothed) residuals within the white square in panel (a) are shown in panel (c). Four bright blobs (A,B,C and D) become visible, which are arranged in a typical strong lensing configuration around the center of the galaxy (G). \emph{Panel (d):} Stellar Kinematics along the galaxy's major axis as measured by \citet{SWELLSV}, line-of-sight rotation velocity $v_\text{rot}$, line-of-sight velocity dispersion $\sigma$ and the rms-velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$. The dotted line in panel (b) indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6" = 5.2$ kpc. The $v_\text{rot}$ curve reveals that J1331 is counter-rotating within $R_\text{eff}$. TO DO: Add (x,y) axis in figure b).???????}
\label{fig:specialJ1331}
\end{figure*}

>>>>

\paragraph{Characteristics of J1331}
\begin{itemize}
\item SDSS J1331+3638 (J1331)
\item approximate hubble type Sb
\item first discovered by Sloan digital sky survey (SDSS) [TO DO: REF]
\item at redshift $z_d\simeq0.113$ [TO DO: REF]
\item \citet{SWELLSI} identified it as a strong lens
\item large reddish bulge and bluish spiral arms, see Fig. \ref{fig:F450W} and \ref{fig:F814W}
\item superimposed by quadruplet of extended bluish images at a redshift of $z_s\simeq0.254$ [TO DO: REF], see Fig. \ref{fig:lens_just_imgpos}
\item lensed object might be a star-forming blob of a background galaxy.
\item Lensing properties first analysed by \citet{SWELLSIII}
\item rather edge-on $\longrightarrow$ possible to measure rotation curves. \citet{SWELLSV} measured the gas and stellar rotation curves along the major axis. Fitted galaxy model to gas kinematics at large radii, and lensing result
\item large counter-rotating core, see Fig. \ref{fig:kinematics}
\item possible minor merger in the past
\end{itemize}

@@

\paragraph{Data used}
\begin{itemize}
\item Hubble Space Telescope (HST)/WFPC2/WFC3 imaging by \cite{SWELLSI}, see Fig. \ref{fig:F450W} and \ref{fig:F814W}
\item  \citet{SWELLSV} measured the gas and stellar rotation curves along the major axis. see Fig. \ref{fig:F814W} and \ref{fig:kinematics}
\end{itemize}

@@

\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in F450W ("blue")}
  \label{fig:F450W}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in F814W ("red")}
  \label{fig:F814W}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{The lensing images}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar Kinematics by \citet{SWELLSV}}
  \label{fig:kinematics}
\end{subfigure}
\caption{Hubble Space telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3638 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST/WFPC2/WFC3 images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The galaxy's coordinates on the sky are right ascension $\alpha$ = 202.91800$^\circ$ and declination $\delta$ = 36.46999$^\circ$ (epoch J2000). Image orientation and scaling are indicated in panel (a); the scaling transformation from arcseconds to the physical size of the galaxy in kpc uses the galaxy's redshift $z_d = 0.113$ \citep{SWELLSIII} (i.e. assumes an angular diameter distance of 414 Mpc). The color scaling of these two images is the same. The black solid line in panel (a) shows the orientation of the major-axis. The line has a length of 10 arcsec and indicates the region within which we carry out the Jeans modelling. [TO DO: NOT ALL THE TIME.???????] \emph{Panel (c):} The central region of J1331 in F450W, surface brightness subtracted. An IRAF ellipse ???? fit to the F450W surface brightness in panel (a) was subtracted from the image. The (smoothed) residuals within the white square in panel (a) are shown in panel (c). Four bright blobs (A,B,C and D) become visible, which are arranged in a typical strong lensing configuration around the center of the galaxy (G). \emph{Panel (d):} Stellar Kinematics along the galaxy's major axis as measured by \citet{SWELLSV}, line-of-sight rotation velocity $v_\text{rot}$, line-of-sight velocity dispersion $\sigma$ and the rms-velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$. The dotted line in panel (b) indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6" = 5.2$ kpc. The $v_\text{rot}$ curve reveals that J1331 is counter-rotating within $R_\text{eff}$. [TO DO: Add (x,y) axis in figure b).???????]}
\label{fig:specialJ1331}
\end{figure*}

### 2015-08-05 docs/model_MGE_v1.01.tex
Started writing sections on MGE photometry.

@@

>>>>

Multi-Gaussian Expansions (MGE) are used to parametrize the observed surface brightness or projected total mass of a galaxy as a sum of $N$ two-dimensional, elliptical Gaussians \citep{1991ApJ...366..599B,1992A&A...253..366M,1994A&A...285..723E,1999MNRAS.303..495E}. This work makes use of the algorithm and code\footnote{Michele Cappellari's IDL code package for fitting MGEs to images is available online at \url{http://www-astro.physics.ox.ac.uk/~mxc/software}. The version from June 2012 was used in this work.} by \citet{Cap02}. We assume all Gaussians to have the same center and position angle $\phi$, i.e. orientation of w.r.t. the $y'$-axis of the coordinate system with polar coordinates $(R',\theta')$ [TO DO: CHECK]. Then the surface brightness can be written as
\begin{eqnarray}
I(R',\theta') &=& \sum_{i=1}^{N} I_{0,i} \exp\left[ - \frac{1}{2\sigma_i^2} \left(x^{'2} + \frac{y^{'2}}{q_i^{'2}}\right)\right]\label{eq:MGEgeneral}\\
\text{with } I_{0,i} &=& \frac{L_i}{2\pi \sigma_i^2 q'_i}\nonumber\\%\label{eq:centralItotalL}\\
\text{and } x'_i &=& R' \cos(\theta' - \phi)\nonumber\\
y'_i &=& R' \sin(\theta' - \phi),\nonumber
\end{eqnarray}
where $I_{0,i}$ is the central surface brightness of each Gaussian, $L_i$ its total luminosity, $\sigma_i$ its dispersion along the major axis and $q'_i$ the axis ratio between the elliptical Gaussians major and minor axis.
\\We can also expand the telescopes point-spread function (PSF) as a sum of circular Gaussians,
\begin{equation}
\text{PSF}(x,y) = \sum_j \frac{G_j}{2 \pi \delta_j^2} \exp\left[- \frac{1}{2 \delta_j^2} \left(x^2 + y^2 \right)\right], \label{eq:PSFgeneral}
\end{equation}
where $\sum_j G_j = 1$ and $\delta_j$ are in this case the dispersions of the circular PSF Gaussians. In this case the observed surface brightness distribution is a convolution of the intrinsic surface brightness in Eq. (\ref{eq:MGEgeneral}) with the PSF in Eq. (\ref{eq:PSFgeneral}): $(I \ast \text{PSF}) (x',y')$ is then again a sum of Gaussians and can be directly fitted to an image of the galaxy in question.

@@

\paragraph{Other stuff to mention}
\begin{itemize}
\item [TO DO] analytic deprojection into 3D Gaussians
\end{itemize}

### 2015-08-05 docs/results_photometry_v1.01.tex
Started writing sections on MGE photometry.

\subsection{Surface Photometry for J1331 with MGE}

@@

\begin{table*}
\centering
\begin{minipage}[t]{40mm}
\caption{PSF F814W MGE ??????}
\begin{tabular}{ccc}
\hline
$k$ & $G_k$ & $\sigma_k$ [arcsec] \\\hline
1 & 0.184 & 0.038\\
2 & 0.485 & 0.085\\
3 & 0.222 & 0.169\\
4 & 0.109 & 0.487\\\hline
\end{tabular}
\label{tab:PSFMGEF814W}
\end{minipage}
\hspace{10mm}
\begin{minipage}[t]{100mm}
\caption{F814W MGE???}
\begin{tabular}{cccccc}
\hline
 & total luminosity  & surface density & \multicolumn{2}{c}{Gaussian dispersion} & axis ratio\\
$k$  & $L_k$ [counts] & $I_{0,k}$ [$L_\odot$/pc$^2$] & $\sigma_k$ [arcsec] & $\sigma_k$ [kpc] & $q'_k$\\\hline
1  &     9425.96 &      20768.  &  0.051   & 0.103  & 1.00\\
2  &    13173.0 &        3161.2 &  0.178   & 0.358  & 0.76\\
3  &    40235.0 &        1588.2 &  0.503   & 1.008  & 0.58\\
4  &    67755.2 &         502.25&  1.180   & 2.368  & 0.56\\
5  &    203677. &         136.51&  3.891   & 7.805  & 0.57\\\hline
\end{tabular}
\label{tab:MGEF814W}
\end{minipage}
\end{table*}

>>>>

\subsection{Surface Photometry for J1331 with MGEs}

To derive J1331's surface brightness distribution, we use the HST image in the infrared F814W filter, shown in Fig. \ref{fig:F814W}. In infrared J1331's central old and smooth stellar component is more extended than in the F450W filter in Fig. \ref{fig:F450W}, which is more sensitive to young clumpy star-forming regions. In infrared the bulge is also much brighter than the bluish lens images and the imaging is less prone to extinction. The F814W image is therefore best suited for fitting a MGE.

\paragraph{PSF for the HST/F814W filter.} We find a radial profile for the HST/F814W filter PSF from circular annuli within a synthetic PSF image from [TO DO: Where did we get this image from???], ignoring diffraction spikes. The one-dimensional MGE fit of Eq. (\ref{eq:PSFgeneral}) to this radial profile is performed using the code by [TO DO: REF and footnote to code]. The MGE parameters of the normalized PSF model are given in Tab. \ref{tab:PSFMGEF814W}.

@@

\begin{table*}
\centering
\begin{minipage}[t]{40mm}
\caption{F814W PSF MGE: Parameters of the circular four-Gaussian MGE in Eq. (\ref{eq:PSFgeneral} fitted to the radial profile of the synthetic HST/F814W PSF image by [TO DO].}
\begin{tabular}{ccc}
\hline
$k$ & $G_k$ & $\delta_k$ [arcsec] \\\hline
1 & 0.184 & 0.038\\
2 & 0.485 & 0.085\\
3 & 0.222 & 0.169\\
4 & 0.109 & 0.487\\\hline
\end{tabular}
\label{tab:PSFMGEF814W}
\end{minipage}
\hspace{10mm}
\begin{minipage}[t]{100mm}
\caption{F814W MGE???}
\begin{tabular}{cccccc}
\hline
 & total luminosity  & surface density & \multicolumn{2}{c}{Gaussian dispersion} & axis ratio\\
$k$  & $L_k$ [counts] & $I_{0,k}$ [$L_\odot$/pc$^2$] & $\sigma_k$ [arcsec] & $\sigma_k$ [kpc] & $q'_k$\\\hline
1  &     9425.96 &      20768.  &  0.051   & 0.103  & 1.00\\
2  &    13173.0 &        3161.2 &  0.178   & 0.358  & 0.76\\
3  &    40235.0 &        1588.2 &  0.503   & 1.008  & 0.58\\
4  &    67755.2 &         502.25&  1.180   & 2.368  & 0.56\\
5  &    203677. &         136.51&  3.891   & 7.805  & 0.57\\\hline
\end{tabular}
\label{tab:MGEF814W}
\end{minipage}
\end{table*}
